---
title: "Script Repo"
output: html_document
---

This is a sample collection from the course Data Science Essentials

### Get dataset from csv
```{r}
filePath <- file.path('./Data', 'Automobile price data _Raw_.csv')
auto.price <- read.csv(filePath, header = TRUE, stringsAsFactors = FALSE)
str(auto.price)
```

### Convert to NAs and remove them
```{r}
## Grab columns to change values of
cols = c('price', 'bore', 'stroke', 'horsepower', 'peak.rpm')
## Convert the character '?' to an NA
auto.price[, cols] = lapply(auto.price[, cols], function(x) ifelse(x == '?', NA, x)) 
## Remove rows with NAs
auto.price = auto.price[complete.cases(auto.price), ] 
```
### Convert character columns to numeric
```{r}
## Convert character columns to numeric
auto.price[, cols] = lapply(auto.price[, cols], as.numeric) 
```
### Transform column
```{r}
auto.price$lnprice = log2(auto.price$price)
```

### Describe function for making summary of column but with standard deviation included
```{r}
## Use the function describe to get extended summary of column
describe = function(df, col){
  tmp = df[, col]
  sumry = summary(tmp)
  nms = names(sumry)
  nms = c(nms, 'std')
  out = c(sumry, sd(tmp))
  names(out) = nms
  out
}

describe(auto.price, 'price')
```

### Plot function for plotting histogram and simple boxplot on certain column
```{r}
plotstats <- function(df, col, bins = 30){
   require(ggplot2)
   require(gridExtra)
   dat = as.factor('') ## Not sure what this does
   
   ## Compute bin width
   bin.width = (max(df[, col]) - min(df[, col]))/ bins

   ## Plot a histogram
   p1 = ggplot(df, aes_string(col)) +
     geom_histogram(binwidth = bin.width)
   
   ## A simple boxplot
   p2 = ggplot(df, aes_string(dat, col)) +
     geom_boxplot() + coord_flip() + ylab('')
   
   ## Now stack the plots
   grid.arrange(p2, p1, nrow = 2)
}

plotstats(auto.price, 'price')
```

### Simple scatter plot function
```{r, echo=FALSE}
## Fcn: Plot two cols against each other
vis.simple <- function(col1, df = auto.price, col2 = 'lnprice'){
  require(ggplot2) # make sure ggplot2 is loaded
  title = paste('Plot of', col1, 'vs.', col2) # create the title text string
  ggplot(df, aes_string(col1, col2)) +
    geom_point() +
    ggtitle(title)
} 

plot.cols <- c("curb.weight",
               "engine.size",
               "city.mpg")

lapply(plot.cols, vis.simple)
```

### Function for creating outlier flags
```{r}
id.outlier <- function(df){
  ## Use ifelse to filter for outliers
  df[, "outlier"] <- ifelse(df[,"engine.size"] > 190 | 
                              df[, "curb.weight"] > 3500| 
                              df[, "city.mpg"] > 40, '1', '0')   
  df
}

## ID the outliers
auto.price <- id.outlier(auto.price)

require(dplyr) # make sure dplyr is loaded
## Return just the outliers with a filter
justOutliers <- auto.price %>% filter(outlier == 1)
## Return without the outlier col and filtered on those that aren't outliers
excludedOutliers <- auto.price %>% filter(outlier == 0) %>% subset(select=-outlier)
```

### Scatter plot function for certain factors (outliers etc)
```{r}
vis.outlier <- function(col1 = 'city.mpg'){
  require(ggplot2) # make sure ggplot2 is loaded
  ## convert character columns to factors for plotting
  auto.price[, "outlier"] <- as.factor(auto.price[, "outlier"])    
  auto.price[, "fuel.type"] <- as.factor(auto.price[, "fuel.type"])    
  title = paste('Plot of', col1, 'vs.lnprice') # character string title
  ggplot(auto.price, aes_string(col1, 'lnprice')) +
    geom_point(aes(color = outlier, 
                   shape = fuel.type, 
				   alpha = 0.5, size = 4)) +
    ggtitle(title)
}

## set the columns to plot
plot.cols <- c("curb.weight",
               "engine.size",
               "city.mpg")

lapply(plot.cols, vis.outlier)  # plot the results

```

### Function for joining dataframes
```{r}
## Define a function to join the data frames
join.auto <- function(autos, makes){
  require(dplyr) ## Make sure dplyr is loaded
  left_join(autos, makes, by = 'make.id')
}

## R code to join the two input tables
autos <- read.csv(file.path('./Data', 'autos.csv'), header=TRUE, stringsAsFactors=FALSE)
makes <- read.csv(file.path('./Data', 'makes.csv'), header=TRUE, stringsAsFactors=FALSE)
str(autos)
str(makes)

out <- join.auto(autos, makes) ## Join the data frames
head(out)
```

### Function to clean and prep the auto data using mutate etc
```{r}
prep.auto <- function(df, col.names){
  require(dplyr) ## Make sure dplyr is loaded
  
  ## set the column names. 
  names(df) <- col.names
  
  ## Eliminate unneeded columns
  df <- df[,!(names(df) %in% 
                c('symboling', 'normalizedlosses', 'makeid'))]
 
  ## Coerce some character columns to numeric 
  ## Uncomment if NOT in Azure ML
  # cols <- c('price', 'bore', 'stroke', 'horsepower', 'rpm')
  # df[, cols] <- lapply(df[, cols], as.numeric)

  ## Add a log transformed column for price using dplyr mutate
  df <- df %>% mutate(lnprice = log(price))
  
  ## Remove rows with NAs 
  df <- df[complete.cases(df), ]
  
  ## Remove duplicate rows
  df <- df %>% filter (! duplicated(df,))
  
  ## Consolidate the number of cylinders
  df <- df %>%
    mutate(cylinders = ifelse(cylinders %in% c("four", "three", "two"), "three-four",
           ifelse(cylinders %in% c("five", "six"), "five-six", "eight-twelve")))
  df
}

## Define column names
col.names <- c('symboling', 'normalizedlosses', 'makeid', 'fueltype', 'aspiration', 'doors',
               'body', 'drive', 'engineloc', 'wheelbase',
               'length', 'width', 'height', 'weight', 'enginetype',
               'cylinders', 'enginesize', 'fuelsystem', 'bore', 'stroke',
               'compression', 'horsepower', 'rpm', 'citympg',
               'highwaympg', 'price', 'make')

## R code to prep the auto data
out <- prep.auto(autos.price, col.names)
str(out)
```

### Using scale function
```{r}
num.cols <- c("wheel.base",
				"length",
				"width",
				"curb.weight")

out <- scale(auto.price[, num.cols])
head(out)
```

###
```{r}

```